Syntax-Based Concept Alignment for Domain-Specific Machine Translation

### 2021-05-09 paper/paper.tex
general part of ce eval

\subsection{Evaluating CE} % TODO: find better title?
% metrics & methods
\subsubsection{Results}
% PUD (incl. reason-wise?): independent, comparison with SMT
% RAW
\subsubsection{Discussion}

>>>>

\subsection{Evaluating CE} %TODO: better title?
While in this work we focus mostly on the MT applications of CA, automatic translation, and much less GF-based domain-specific tranlation, is not the only context in which our system can be put to use. 
For instance, it is easy to imagine using it to improve the UI of online translation memories or to facilitate reading parallel texts in the context of language learning.
For this reason, a first set of experiments is aimed at evaluating the alignments obtained with our CA module independently from the subsequent stages of our proposed lexicon generation pipeline.

We assess the correctness of the alignments the CE module is able to extract first from manually annotated data, then from the DMI and CSE corpus, comparing our results with those obtained with a standard statistical tool, \texttt{fast\_align} \cite{TODO:}.
Doing so requires manually labelling the correspondences as either correct or incorrect.

For measuring the performance of our system, while precision and recall are two well-known metrics that would be very well suited to the task, the lack of a gold standard for CE imposes us to approximate them with respectively:

\begin{itemize}
  \item the number of correct alignments the program is able to extract
  \item the ratio between such number and the total number of alignments extracted.
 \end{itemize}

However, since some alignments are only correct in the specific context of the sentence pair in which they occur, we make a further distinction between correct alignments that are relevant for a translation lexicon, marked as \texttt{+} in Section \ref{ceres}'s tables, and alignments that are useful for comparing the sentences but should not be used for MT, marked as \texttt{=} instead. % TODO: boat/train example 
Incorrect correspondences are marked as \texttt{-}.

\subsubsection{Results} \label{ceres}
% PUD (incl. reason-wise?): independent, comparison with SMT
% RAW
\subsubsection{Discussion}
